### 2015-08-03 implementation-specs.tex

\subsubsection{Configuration Complexity}
A spec DAG has many degrees of freedom, and iusers cannot reasonably be 
expected to understand or to specify all of them.  In our experience at LLNL, 
the typical user only cares about a small number of build constraints (if any),
and does not know enough to
specify the rest. For example, a user may know that a certain version of a library like
{\tt boost}~\cite{boost} is required, but only cares that other build parameters are set so that
the build will succeed.
%
Configuration complexity makes the HPC software ecosystem difficult to manage: 
too many parameters exist to know all of them. However, the known and 
important ones often provide detailed build constraints. Thus, we have two 
competing concerns.  We need the ability to specify details
without having to remember all of them.

@@

We have developed a syntax for specs that allows users to specify
only the constraints that matter. Our syntax can represent software 
DAGs concisely enough to use it on the command 
line. The spec syntax is recursively defined to allow users to specify
parameters on dependencies as well as on the root of the DAG. 
Figure~\ref{fig:grammar} showsh the EBNF grammar through which Spack 
flexibly supports constraints.

@@

Figure~\ref{fig:specs-mpileaks-abstract} shows the recursive nature of the spec syntax:
%
\begin{minted}[fontsize=\scriptsize]{bash}
    $ spack install mpileaks@2.3 ^callpath@1.0+debug ^libelf@0.8.11
\end{minted}
%
The caret (\verb|^|) denotes constraints for a particular dependency. The DAG,
now has version constraints on {\tt callpath} {\it and} {\tt libelf},
and the user has requested the {\tt debug} variant of the {\tt callpath} library.

Recall that Spack guarantees that only a single version of any package occurs 
in a spec.  Within a spec, each dependency can be uniquely
identified by its package name alone.  Therefore, the user does not need to consider
DAG connectivity to add constraints but instead only must only know that 
{\tt mpileaks} depends on {\tt callpath}.
Thus, dependency constraints can appear in an arbitrary order.

@@

{\bf Cross-compilation.}
To support cross-compilation, spec can include the system architecture 
(line 5). Platforms begin with {\tt =} and take names like {\tt linux-ppc64} 
or {\tt bgq}.  They are specified per-package. This mechanism allows front-end 
tools to depend on their back-end measurement
libraries with a {\it different} architecture on cross-compiled machines.

>>>>

\subsubsection{Configuration Complexity}
A spec DAG has many degrees of freedom, and users cannot reasonably be
expected to understand or to specify all of them.  In our experience at LLNL,
the typical user only cares about a small number of build constraints (if any),
and does not know enough to
specify the rest. For example, a user may know that a certain version of a library like
{\tt boost}~\cite{boost} is required, but only cares that other build parameters are set so that
the build will succeed.
%
Configuration complexity makes the HPC software ecosystem difficult to manage:
too many parameters exist to know all of them. However, the known and
important ones often provide detailed build constraints. Thus, we have two
competing concerns.  We need the ability to specify details
without having to remember all of them.

@@

We have developed a syntax for specs that allows users to specify
only the constraints that matter. Our syntax can represent software
DAGs concisely enough to use it on the command
line. The spec syntax is recursively defined to allow users to specify
parameters on dependencies as well as on the root of the DAG.
Figure~\ref{fig:grammar} shows the EBNF grammar through which Spack
flexibly supports constraints.

@@

Figure~\ref{fig:specs-mpileaks-abstract} shows the recursive nature of the spec syntax:
%
\begin{minted}[fontsize=\scriptsize]{bash}
    $ spack install mpileaks@2.3 ^callpath@1.0+debug ^libelf@0.8.11
\end{minted}
%
The caret (\verb|^|) denotes constraints for a particular dependency. The DAG
now has version constraints on {\tt callpath} {\it and} {\tt libelf},
and the user has requested the {\tt debug} variant of the {\tt callpath} library.

Recall that Spack guarantees that only a single version of any package occurs
in a spec.  Within a spec, each dependency can be uniquely
identified by its package name alone.  Therefore, the user does not need to consider
DAG connectivity to add constraints but instead must only know that
{\tt mpileaks} depends on {\tt callpath}.
Thus, dependency constraints can appear in an arbitrary order.

@@

{\bf Cross-compilation.}
To support cross-compilation, specs can include the system architecture
(line 5). Platforms begin with {\tt =} and take names like {\tt linux-ppc64}
or {\tt bgq}.  They are specified per-package. This mechanism allows front-end
tools to depend on their back-end measurement
libraries with a {\it different} architecture on cross-compiled machines.
